---
title: "French given names per year per department"
author: "Raised by Lucas Mello Schnorr, Jean-Marc Vincent; reviewed by Gregory James"
date: "October, 2023"
output:
  pdf_document: 
    df_print: paged
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# The problem context
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time.

[https://www.insee.fr/fr/statistiques/2540004](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2021_csv.zip), we choose this dataset because it is sufficiently large, you can't do the analysis by hand, the structure is simple


You need to use the _tidyverse_ for this analysis. Unzip the file _dpt2020_txt.zip_ (to get the **dpt2020.csv**). Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

## Download Raw Data from the website
```{r}
file = "dpt2021_csv.zip"
if(!file.exists(file)){
  download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2021_csv.zip",
	destfile=file)
}
unzip(file)
```
Check if your file is the same as in the first analysis (reproducibility)
```{bash eval=FALSE, include=FALSE}
md5 dpt2021.csv
```
expected :
MD5 (dpt2021.csv) = f18a7d627883a0b248a0d59374f3bab7

## Build the Dataframe from file

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
df_loaded <- read_delim("dpt2021.csv",delim=";",show_col_types = FALSE)
```

All of these following questions may need a preliminary analysis of the data, feel free to present answers and justifications in your own order and structure your report as it should be for a scientific report.

1. Choose a firstname and analyse its frequency along time. Compare several firstnames frequency
2. Establish by gender the most given firstname by year. Analyse the evolution of the most frequent firstname.
3. Optional : Which department has a larger variety of names along time ? Is there some sort of geographical correlation with the data?


-------------------------------------------------

-------------------------------------------------

## Cleaning the Dataframe

The preliminary step before running some analysis on the evolution of first names in France, we have to first analyze and ensure that the database is structured in the right way with necessary cleaning performed. We are therefore first printing different details on the "dataframe loaded"
```{r warning=FALSE}
print(head(df_loaded))
print(tail(df_loaded))
dim(df_loaded)
length(df_loaded)
summary(df_loaded)
```

The first table is showing the value "_PRENOMS_RARES" as the firstname which we will investigate after cleaning some other fields.

### Titles of the Table

Titles are not really explicit and I have made the decision to rename those as per below:
  - "sexe" changed to "gender"
  - "preusuel" changed to "firstname"
  - "annais changed to "year"  
  - "dpt" changed to "local_department" 
  - "nombre" changed to "count_of_name"

In order to keep a trace of the original load I have created a new table called "df_loaded_cleaned"

```{r}
df_loaded_cleaned <- df_loaded %>% rename(gender = sexe, firstname = preusuel, year =  annais,
                                    local_department = dpt, count_of_name = nombre)
print(tail(df_loaded_cleaned))
```

### Understanding and Cleaning "XXXX" and "XX" Values

It appears also that there are some data that need to be cleaned out of the table.
I have decided to start investigating the "XXXX" and "XX" values.
To further the investigation, I have looked into the detail for 3 names that are shown with their values.
I am also testing if there are any combination for which we have "XXXX" in year and/or "XX" in local_department.

```{r}
df_count_names <- df_loaded_cleaned %>% count(firstname, sort = TRUE, name = "rows_per_name")
set.seed(1)
name_tested_A <- df_count_names %>% filter(rows_per_name == 6) %>% 
  sample_n(1, replace = FALSE) %>% select(1) %>% as.character()
set.seed(1)
name_tested_B <- df_count_names %>% filter(rows_per_name == 7) %>% 
  sample_n(1, replace = FALSE) %>% select(1) %>% as.character()
set.seed(1)
name_tested_C <- df_count_names %>% filter(rows_per_name == 8) %>% 
  sample_n(1, replace = FALSE) %>% select(1) %>% as.character()

count(df_loaded_cleaned, year == "XXXX" & local_department != "XX")
count(df_loaded_cleaned, year != "XXXX" & local_department == "XX")
count(df_loaded_cleaned, year != "XXXX" & local_department != "XX")
count(df_loaded_cleaned, year == "XXXX" & local_department == "XX")

print(df_loaded_cleaned %>% filter(firstname == name_tested_A))
print(df_loaded_cleaned %>% filter(firstname == name_tested_B))
print(df_loaded_cleaned %>% filter(firstname == name_tested_C))
```

The outcome of that investigation is that any logical information can established but an assumption can be made.
I have considered that those records correspond, for a first name, to all the records for which the year and/or the department haven't been recorded.
To avoid removing a data that might be useful in the analysis I have decided to replace "XXXX","XX" by "9999".

```{r message=TRUE, warning=FALSE, paged.print=TRUE}
df_loaded_cleaned$year[df_loaded_cleaned$year == "XXXX"] <- "9999"
df_loaded_cleaned$local_department[df_loaded_cleaned$local_department == "XX"] <- "9999"
print(df_loaded_cleaned %>% filter(firstname == name_tested_A))
```

### Format of the Data

In addition I have also seen that the column that includes the year information is shown as "chr" which might generate wrong analysis if we want to use that base and compare to an evolution over time.
I am therefore changing this column to become an integer ("int").

```{r}
set.seed(7)
df_loaded_cleaned$year <- as.numeric(df_loaded_cleaned$year)
names_tested_A <- df_count_names %>% filter(rows_per_name == 7) %>% 
  sample_n(1, replace = FALSE) %>% select(1) %>% as.character()
print(df_loaded_cleaned %>% filter(firstname == names_tested_A))
any(is.na(df_loaded_cleaned))
which(is.na(df_loaded_cleaned))
```

It appears that there are some NA values in the table. I am therefore looking in every columns to see where there are some and identify the action to take with those values.

```{r warning=FALSE, paged.print=TRUE}
paste0("Test NA in column 'gender' =",any(is.na(df_loaded_cleaned$gender)))
paste0("Test NA in column 'firstname' =",any(is.na(df_loaded_cleaned$firstname)))
paste0("Test NA in column 'year' =",any(is.na(df_loaded_cleaned$year)))
paste0("Test NA in column 'local_department' =",any(is.na(df_loaded_cleaned$local_department)))
paste0("Test NA in column 'count_of_name' =",any(is.na(df_loaded_cleaned$count_of_name)))
```



```{r}
count(df_loaded_cleaned, is.na(df_loaded_cleaned$firstname))
which(is.na(df_loaded_cleaned$firstname))
```
```{r}
df_loaded_cleaned[3323341,]
```

Only one firstname is shown as NA and looking at the information it seems that there are some records for which names are missing.
I am again updating this value with "9999" to make sure that we can then use the base appropriately.

```{r}
df_loaded_cleaned[3323341,2] <- "9999"
df_loaded_cleaned[3323341,]
```

```{r}
print(head(df_loaded_cleaned))
print(tail(df_loaded_cleaned))
dim(df_loaded_cleaned)
length(df_loaded_cleaned)
summary(df_loaded_cleaned)
```


### Rare First Names

When running the first review of the table it appears that the name "_PRENOMS_RARES" is shown as the first name.
I have excluded this value and run a summary to see if it is linked to the count of names.

```{r}
firstname_count_mapping <- df_loaded_cleaned %>% select(!(gender)) %>% select(!(year)) %>% 
  select(!(local_department))
firstname_count_mapping_exc_rare <- firstname_count_mapping[firstname_count_mapping$firstname != 
                                                              "_PRENOMS_RARES", ]
summary(na.omit(firstname_count_mapping_exc_rare))
```

We can clearly see that first names that are registered are the one that have more than 2 records in the given range (gender, year, department). All the others are registered as "_PRENOMS_RARES".
To be sure that we are having all the data and numbers available for the analysis i have decided to keep the data as it but can still remove this value when needed.

After all those changes the inital dataframe structured is preserved but the data is in a format that can be used for making further analysis on the repartition of first names.

## Analysis over the first name frequency along time. 

For this analysis the local_department nor the gender information are needed. As a starting point I have created a new table and summed up all the count of names per year and remove the local_department and gender_data.

```{r}
table_freq_analysis <- df_loaded_cleaned %>% select(!(gender)) %>% select(!(local_department)) %>%
  group_by(firstname,year,.add = TRUE,.drop = FALSE) %>% 
  mutate(sum_per_year = sum(count_of_name)) %>% select(!(count_of_name))
table_freq_analysis <- table_freq_analysis[!duplicated(table_freq_analysis),]
summary(table_freq_analysis)
```

As we are going to look at the distribution per year I have also removed the data for which the year is '9999'

```{r}
table_freq_analysis <- table_freq_analysis[table_freq_analysis$year != 9999, ]
summary(table_freq_analysis)
```


```{r}
filter_table_name = function (name, tabx){
  FirstNameFiltered <- tabx %>% filter(firstname==name)
}
```

```{r warning=FALSE}
set.seed(6)
randomrows <- sample(nrow(table_freq_analysis))
table_freq_analysis_random <- table_freq_analysis[randomrows, ]
list_firstnames_random <- unique(table_freq_analysis_random[,1])

table_freq_analysis_names <- table_freq_analysis[0,]
for (each_rows in sample(5)){
  names_selected_x <- list_firstnames_random[each_rows,1] %>% as.character()
  table_freq_analysis_names <- table_freq_analysis_names %>% 
    rbind(filter_table_name(names_selected_x,table_freq_analysis))
  }

table_freq_analysis_name_1 <- table_freq_analysis_names %>% 
  filter(firstname == list_firstnames_random[1,1]) 

plot(table_freq_analysis_name_1$year,table_freq_analysis_name_1$sum_per_year, 
     xlab="Year", ylab="Count", main = paste0("Distribution of the First Name '", 
                   list_firstnames_random[1,1],"' per Year"))
```

A random first name "JOFFREY" is selected and a graph shows, using a plot, how that first name has evolved over time with a pic around 1990.

I have also made a random selection of 5 names and run the same graph.

```{r}
ggplot(data = table_freq_analysis_names, aes(x=year, y=sum_per_year, color=firstname)) + 
  geom_point() + ggtitle("Distribution of a Random Selection of First Name per Year")
```

This type of graph allows to see that some first name are "stars" during some periods ("DANIEL" have been given a lot around 1950).
We can also see that some are appearing/diseappiring along time. 
We still need to find another way to represent the data as when a name is having a big number, it is compressing all the other data.
This is especially the around 1950's, due to baby boom, during when a much higher number of first names have been given.

### Compare several first names frequency

In order to get a smoother vision over the data, I have added the frequency of a first name overtime.
We can again plot the data to see the result.

```{r}
table_freq_analysis <- table_freq_analysis %>% group_by(firstname) %>% 
  mutate(sum_per_name = sum(sum_per_year)) %>% 
  mutate(frequency_per_year = sum_per_year/sum_per_name)

set.seed(6)
randomrows <- sample(nrow(table_freq_analysis))
table_freq_analysis_random <- table_freq_analysis[randomrows, ]
list_firstnames_random <- unique(table_freq_analysis_random[,1])

table_freq_analysis_names <- table_freq_analysis[0,]
for (each_rows in sample(5)){
  names_selected_x <- list_firstnames_random[each_rows,1] %>% as.character()
  table_freq_analysis_names <- table_freq_analysis_names %>% 
    rbind(filter_table_name(names_selected_x,table_freq_analysis))
  }
ggplot(data = table_freq_analysis_names, aes(x=year, y=frequency_per_year, color=firstname)) + 
  geom_point() + ggtitle("Normalized Distribution of a Random Selection of First Name per Year")
```

We can see more details on each names and easily identify when they had their "star" moments.

We can also look at the first name that have been given more than 150 times in a year and where there is the higher frequence to have a look at the first names that had their "star moments" each year.

```{r}
table_max_freq_analysis_names <- table_freq_analysis %>% filter(sum_per_year > 150) %>% 
  group_by(year) %>% top_n(1, frequency_per_year)
ggplot(subset(table_max_freq_analysis_names, year >= 1970 & year <= 2021), 
       aes(x=year, y=frequency_per_year,label = firstname)) + 
  ggtitle("First Name per Year with highest individual frequence per year")+
  geom_text(size = 2,check_overlap = FALSE,position = position_stack(vjust = 0.5))+
  facet_wrap(~year)+theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  theme(axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank())
```


### Most Given First Name by Gender

We are now trying to find the most given first name by gender.
The exact same steps done before are used without removing the gender information.
The value with "_PRENOMS_RARES" will also impact the result so I am removing it.
In addition, the table with the first name with the first name most given by gender in a year.

```{r}
table_most_analysis <- df_loaded_cleaned %>% select(!(local_department)) %>%
  group_by(gender,firstname,year,.add = TRUE,.drop = FALSE) %>% 
  mutate(sum_per_year = sum(count_of_name)) %>% select(!(count_of_name))
table_most_analysis <- table_most_analysis[!duplicated(table_most_analysis),]
table_most_analysis <- table_most_analysis[table_most_analysis$year != 9999, ]
table_most_analysis <- table_most_analysis[table_most_analysis$firstname != "_PRENOMS_RARES", ]
table_most_analysis <- table_most_analysis %>% ungroup() %>% 
  group_by(gender,year) %>% top_n(1, sum_per_year)
print(tail(table_most_analysis))
```

```{r}
ggplot(subset(table_most_analysis, gender == 1), aes(x=gender, y=gender,label = firstname)) + 
  geom_text(size = 2,check_overlap = FALSE,position = position_stack(vjust = 0.5))+
  ggtitle("Female: Most Given First Names per Year")+
  facet_wrap(~year)+theme(axis.title.x=element_blank()
                          ,axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  theme(axis.title.y=element_blank(),axis.text.y=element_blank()
        ,axis.ticks.y=element_blank(),strip.text = element_text(size=6) )
ggplot(subset(table_most_analysis, gender == 2), aes(x=gender, y=gender,label = firstname)) + 
  ggtitle("Male: Most Given First Names per Year")+
  geom_text(size = 2,check_overlap = FALSE,position = position_stack(vjust = 0.5))+
  facet_wrap(~year)+theme(axis.title.x=element_blank()
                          ,axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  theme(axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()
        ,strip.text = element_text(size=6) )
```

We can see that there are some trends where the top first name is rarely at the top only one year. This is even more the case in the early of the 20th century when "JEAN" and "MARIE" have been given a lot during that period.

```{r}
ggplot(subset(table_most_analysis), aes(x=firstname, fill=gender)) + 
  geom_bar() + ggtitle("Most Given First Names Over Time by Gender")+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 6))
ggplot(subset(table_most_analysis, gender == 1), aes(x=firstname)) + 
  geom_bar() + ggtitle("Male: Most Given First Names Over Time")+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
ggplot(subset(table_most_analysis, gender == 2), aes(x=firstname)) + 
  geom_bar() + ggtitle("Female: Most Given First Names Over Time")+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

When grouping by first name we can see that there are not too many variations in the list the "most" given names.

### Department With Largest Variety of Names

In order to look at how names are distributed within each departments we will again rebuilt a table using the same logic without the "99" and "9999" data.
We are removing the count of each names and adding a new columns with:
  - total number of names given (number of birth)
  - number of different first names per year and department
We can keep the "_PRENOMS_RARES" as it will be counted once only per department and will add information to the dataset.

We are searching which departments have had the largest variety over time by looking at the uniquess disctribution of first names over time.

```{r}
table_by_dpt_analysis <- df_loaded_cleaned %>% select(!(gender)) %>%
  group_by(local_department,year,.add = TRUE,.drop = FALSE) %>% 
   mutate(unique_names_per_department = n()) %>% 
   mutate(total_birth_per_department = sum(count_of_name)) %>% 
  select(!(firstname)) %>% select(!(count_of_name))
table_by_dpt_analysis <- table_by_dpt_analysis[!duplicated(table_by_dpt_analysis),]
table_by_dpt_analysis <- table_by_dpt_analysis %>% ungroup() %>% 
  group_by(local_department,.add = TRUE,.drop = FALSE) %>% 
   select(local_department , total_birth_per_department , unique_names_per_department) %>% 
  summarize(iterations = n(),mean_unique_names_per_department = mean(unique_names_per_department), 
            sd_unique_names_per_department = sd(unique_names_per_department))
table_by_dpt_analysis <- table_by_dpt_analysis[table_by_dpt_analysis$local_department != 9999, ]
print(table_by_dpt_analysis)
```


```{r}
ggplot(data = table_by_dpt_analysis, aes(x=local_department, y=mean_unique_names_per_department)) + 
  geom_col() + scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) + 
  ggtitle("Variety of Names Per Departments Over Time: Mean Number of Unique Names")
```

We can seems some variations over the number of department which could be also linked to where the biggest number of birth have happened.

```{r}
table_by_dpt_analysis_top20 <- table_by_dpt_analysis %>% 
  select(local_department,mean_unique_names_per_department)
table_by_dpt_analysis_top20 <- table_by_dpt_analysis_top20[!duplicated(table_by_dpt_analysis_top20),]
table_by_dpt_analysis_top20 <- table_by_dpt_analysis_top20 %>% 
  top_n(20, mean_unique_names_per_department)
ggplot(data = subset(table_by_dpt_analysis, local_department %in% 
                       table_by_dpt_analysis_top20$local_department), 
       aes(x=local_department, y=mean_unique_names_per_department)) + 
  geom_col() + ggtitle("Variety of Names Per Departments Over Time: Mean Number of Unique Names
                       (Top 20)")
```

It is confirmed that the departments where the biggest average/mean numbers of birth have happened over time and we can search for more explanation by comparing with the number of birth.
This will let us know how variable a given name is for each departments.
